Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning

### 2018-12-01 documentation/paper.tex
Lots of clarifications of various things

@@

Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the most effective region of the hyperparameter search space (L2 $\lambda$ of 0.003 and 2 layers for $\beta_8$ input, and L2 $\lambda$ of 0.001 and dropout of 0 for $\beta_{50,001}$ input). It is very clear that higher resolution data does not improve accuracy in a relatively shallow network.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of {\it FourierMLP}$(\beta_{N})$ configurations in the best-performing region of the hyperparameter search space.} \label{fft_wall_cuts}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            Input Format & L2 $\lambda$ & Dropout & Layers & Accuracy & Precision & Recall & CWSD \\
            \hline
            $\beta_8$ & 0.003 & 0 & 2 & 96.2\% & 94.1\% & 99.8\% & 0.28 \\
            \hline
            $\beta_8$ & 0.003 & 0.25 & 2 & 96.7\% & 94.9\% & 99.5\% & 0.27 \\
            \hline
            $\beta_8$ & 0.003 & 0.5 & 2 & 96.7\% & 95.0\% & 99.9\% & 0.30 \\
            \hline
            $\beta_{50,001}$ & 0.001 & 0 & 2 & 89.1\% & 84.0\% & 100.0\% & 0.47 \\
            \hline
            $\beta_{50,001}$ & 0.001 & 0 & 3 & 92.2\% & 88.8\% & 100.0\% & 0.44 \\
            \hline
            $\beta_{50,001}$ & 0.001 & 0 & 4 & 93.0\% & 89.8\% & 100.0\% & 0.40 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}
% cat ../experimental_data/training_logs/2018_10_26_banded_ap_similarity_last_300_epochs.out | grep examples256 | grep lambda0.003 | ./mean_ap_accuracy_across_configurations.py
% cat ../experimental_data/training_logs/2018_11_24_high_resolution_ap_similarity_last_epoch.out | grep dropout0/ | grep l2_lambda0.001 | ./mean_ap_accuracy_across_configurations.py

@@

Table \ref{super_overview} compares the mean accuracy, precision, recall, and class-wise standard deviation values of each of the input formats.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Comparison of supervised learning input formats.} \label{super_overview}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Configuration & Accuracy & Precision & Recall & CWSD \\
            \hline
            {\it DeepConv}$(\omega)$ & 87.5\% & 71.9\% & 100.0\% & 0.55 \\
            \hline
            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 96.5\% & 94.6\% & 99.7\% & 0.28 \\
            \hline
            {\it FourierMLP}$(\beta_{50,001})$ & 91.4\% & 87.8\% & 100.0\% & 0.44 \\
            \hline
            {\it ImageConv}$(\iota)$ & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

The key conclusion to be drawn from the above table is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format. Higher resolutions seem to make performance worse, likely due to overfitting.

Based on this result, experimentation with semi-supervised learning proceeded using \\ {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$, applying a three-layer perceptron architecture. There were three main reasons for this:

\begin{enumerate}
    \item This was a highly effective configuration, producing 96\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
    \item The use of the exact same input format as AP makes the algorithms more directly comparable.
    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because there are different position distributions of the two classes (as described in Section \ref{og_analysis}).
\end{enumerate}

>>>>

(Side note: intuitively, it would make sense to provide the position $\chi$ directly to the neural network when corrections cannot be applied. However, this cannot presently be done because $\chi$ is a bias for the calibration run type, as described in Section \ref{og_analysis}; neutron sources are placed on one side of the vessel, so neutron events occur more often near that side.)

@@

Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators from the best-performing region of the hyperparameter search space.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of {\it FourierMLP}$(\beta_{N})$ configurations in the best-performing region of the hyperparameter search space.} \label{fft_wall_cuts}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            Input Format & L2 $\lambda$ & Dropout & Layers & Accuracy & Precision & Recall & CWSD \\
            \hline
            $\beta_8$ & 0 & 0.5 & 2 & 98.1\% & 97.1\% & 100.0\% & 0.24 \\
            \hline
            $\beta_8$ & 0.001 & 0.5 & 2 & 96.6\% & 94.7\% & 99.6\% & 0.27 \\
            \hline
            $\beta_8$ & 0.003 & 0.5 & 2 & 96.3\% & 94.3\% & 100.0\% & 0.28 \\
            \hline
            $\beta_{50,001}$ & 0.001 & 0 & 2 & 89.1\% & 84.0\% & 100.0\% & 0.47 \\
            \hline
            $\beta_{50,001}$ & 0.001 & 0 & 3 & 92.2\% & 88.8\% & 100.0\% & 0.44 \\
            \hline
            $\beta_{50,001}$ & 0.001 & 0 & 4 & 93.0\% & 89.8\% & 100.0\% & 0.40 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}
% cat ../experimental_data/training_logs/2018_12_01_banded_all_examples_ap_similarity_last_300_epochs.out | grep dropout0.5 | ./mean_ap_accuracy_across_configurations.py
% cat ../experimental_data/training_logs/2018_11_24_high_resolution_ap_similarity_last_epoch.out | grep dropout0/ | grep l2_lambda0.001 | ./mean_ap_accuracy_across_configurations.py

@@

\subsection{Performance Comparison}

Table \ref{super_overview} compares the mean accuracy, precision, recall, and class-wise standard deviation values (of the highest-performance region of the hyperparameter search space) for each of the input formats.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Comparison of supervised learning input formats.} \label{super_overview}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Configuration & Accuracy & Precision & Recall & CWSD \\
            \hline
            {\it DeepConv}$(\omega)$ & 87.5\% & 71.9\% & 100.0\% & 0.55 \\
            \hline
            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$ & 97.0\% & 95.4\% & 99.9\% & 0.26 \\
            \hline
            {\it FourierMLP}$(\beta_{50,001})$ & 91.4\% & 87.8\% & 100.0\% & 0.44 \\
            \hline
            {\it ImageConv}$(\iota)$ & 63.0\% & 63.0\% & 100.0\% & 1.00 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

The key conclusion to be drawn from the above table is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format. The raw waveform and high-resolution Fourier transform make performance worse. This may be caused by one of the following three reasons:

\begin{itemize}
    \item Insufficient network depth to process higher resolutions
    \item Overfitting on extraneous data
    \item Position corrections are essential to good performance
\end{itemize}

The last reason is very likely. At a fundamental level, $\beta _8$ does not contain any more information than $\beta _{50,001}$, so there is no obvious reason why (with the right network architecture) it should not perform at least as well. While repurposing the position corrections to work on arbitrary-resolution data is outside the scope of this study, it seems a promising subject of future investigation.

Based on this result, experimentation with semi-supervised learning proceeded using \\ {\it FourierMLP}$(${\it PosCor}$(\beta_{8}, \chi))$, applying a three-layer perceptron architecture. There were two main reasons for this:

\begin{enumerate}
    \item This was a highly effective configuration, producing 97.0\% accuracy and a class-wise standard deviation dramatically better than the original PICO-60 neural network on that data set.
    \item If position correction is not applied, there is some risk of introducing a bias to the network's training data because alpha and neutron events have different position distributions (as described in Section \ref{og_analysis}).
\end{enumerate}
